---
title: "Q3"
author: "Deepika"
date: "March 16 2024"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
#knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Loading the dataset

```{r}
str(data)
```
library(dplyr)
dataCountrytop <- data %>%
  group_by(Country) %>%
  summarise(Revenue = sum(Revenue)) %>%
  arrange(desc(Revenue)) %>%
  head(10)
print(dataCountrytop)

library(ggplot2)


plot <- ggplot(dataCountrytop, aes(x = Country, y = Revenue, fill = Revenue)) +
  geom_bar(stat = "identity") +
  labs(title = "Highest Revenue Countries", x = "Country", y = "Revenue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_gradient(low = "blue", high = "red") + coord_flip() 
print(plot)
=======
data$Date <- as.Date(data$Date, format="%Y-%m-%d") 
data$Month <- format(data$Date, "%Y-%m")
data$Quarter <- quarters(data$Date)
data$TransactionValue <- with(data, Quantity * Price) 
monthly_sales <- aggregate(TransactionValue ~ Month, data=data, FUN=sum)
ggplot(monthly_sales, aes(x=Month, y=TransactionValue)) + 
  geom_bar(stat="identity", fill="skyblue") +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Monthly Sales Trends", x="Month", y="Total Sales")
quarterly_sales <- aggregate(TransactionValue ~ Quarter, data=data, FUN=sum)
anova_result <- aov(TransactionValue ~ Quarter, data=data)
summary(anova_result)
ggplot(quarterly_sales, aes(x = Quarter, y =TransactionValue )) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Quarterly Sales Trends", x = "Quarter", y = "Total Sales") +
  theme_minimal()
anova_result <- aov(TransactionValue ~ Quarter, data=data)
summary(anova_result)
```

