```{r}
library(readr)
library(dplyr)
library(ggplot2)

# Read the dataset and suppress column type messages
data <- read_csv("Onlinearticle.csv", show_col_types = FALSE)  

# Calculate average shares for each day of the week directly
average_shares_by_day <- data.frame(
  Day = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),
  Average_Shares = c(
    mean(data$shares[data$weekday_is_monday == 1]),
    mean(data$shares[data$weekday_is_tuesday == 1]),
    mean(data$shares[data$weekday_is_wednesday == 1]),
    mean(data$shares[data$weekday_is_thursday == 1]),
    mean(data$shares[data$weekday_is_friday == 1]),
    mean(data$shares[data$weekday_is_saturday == 1]),
    mean(data$shares[data$weekday_is_sunday == 1])
  )
)

# Plot the results using ggplot2
g <- ggplot(average_shares_by_day, aes(x = Day, y = Average_Shares, fill = Day)) +
  geom_bar(stat = "identity", color = "black") +
  theme_minimal() +
  labs(title = "Average Shares by Day of the Week", x = "Day of the Week", y = "Average Shares") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(g)

```
What impact does publishing time have on the immediate engagement (shares within the first 24 hours) of articles, and how can we optimize publication times based on this data?

Strategic Content Adjustments

Considering articles published over the last year, which have exceeded 5,000 shares, what common characteristics do these articles have, and how can we apply these findings to new content?

What trends can be observed in the sharing behavior over different seasons, and how should we tailor our content calendar to these trends?
```{r}
# Load necessary libraries

# Assume a reference collection date
reference_date <- as.Date("2015-12-31")

# Calculate publication dates by subtracting timedelta from the reference date
data$publication_date <- reference_date - data$timedelta

# Initialize the Season column
data$Season <- character(nrow(data))

# Manually categorize each date into seasons
for (i in 1:nrow(data)) {
  month <- as.integer(format(data$publication_date[i], "%m"))
  if (month %in% c(12, 1, 2)) {
    data$Season[i] <- "Winter"
  } else if (month %in% c(3, 4, 5)) {
    data$Season[i] <- "Spring"
  } else if (month %in% c(6, 7, 8)) {
    data$Season[i] <- "Summer"
  } else {
    data$Season[i] <- "Autumn"
  }
}

# Prepare to calculate average shares for each season
seasons <- unique(data$Season)
seasonal_shares <- numeric(length(seasons))
names(seasonal_shares) <- seasons

# Calculate average shares for each season
for (season in seasons) {
  seasonal_shares[season] <- mean(data$shares[data$Season == season], na.rm = TRUE)
}

# Convert list to data frame for plotting
seasonal_shares_df <- data.frame(Season = names(seasonal_shares), Average_Shares = unlist(seasonal_shares))

# Sort by season for better visualization
season_order <- c("Winter", "Spring", "Summer", "Autumn")
seasonal_shares_df$Season <- factor(seasonal_shares_df$Season, levels = season_order)
seasonal_shares_df <- seasonal_shares_df[order(seasonal_shares_df$Season),]

# Plot the results using base R
barplot(height = seasonal_shares_df$Average_Shares, names.arg = seasonal_shares_df$Season, col = rainbow(length(seasonal_shares_df$Season)),
        main = "Average Shares by Season", xlab = "Season", ylab = "Average Shares", las = 1, cex.names = 0.8)
```

